Evaluating Cultural Transmission in Bronze Age burial rites of Central, Northern and North-western Europe using radiocarbon data

### 2019-05-26 review/answer_to_review_2.md

Two major take aways: (1) Their mobility estimator could be used for a better, region independent statistic for the general degree of interaction. (2) I could as well use a moving window instead of 200 years time slices (Figure 6).

I added some sentences in the conclusion to give an outlook on future work.

> I also still have some issues with how sample size is effectively being ignored once proportions are computed, and I think this should be emphasised more strongly as I expect not everyone would be reading beyond the conclusion section.

I made some major changes to the conclusion to make this very clear.

@@

Rephrased. Added the important "no mutation or innovation" conditions.

> Page 3, Column 1, Lines 2-6. Great!

> Page 3, Column 2, Line 59. ‘the region are circular’ Why this shape if it causes overlap? Why not hexagons?

This is true. I added a sentence to point this out. For this paper I will use the established regions, because changing them at this point would render the area discussions useless. In the future more data might be available and it should be possible to establish an a lot more fine grained, hexagonal or rectangular region grid.

@@

I modified the small, introductory paragraph of this section to make it clear, that this is a thought experiment based on a very small amount of data. I also set the narrative in italics.

> Table 3 & Figure 6. The figure nicely summarises the trend detailed on Table 3, but does not distinguish instances where the correlation were identified as signficant vs those that were not. Indeed in no cases the analysis shows a signficant correlation under the Partial Mantel test, whilst only few chronological blocks are showing signficant correlations. I suggest highlighting this by perhaps changing the the point symbology. It is also worth noting that if anything there is a bias towards type I error given: a) that we are dealing with multiple testing; and b) these analyses do not take into account sample sizes which are often very small. P-values are not the core issue here, but rather the fact that some way to at least visually convey this uncertainty is much needed.

To address this issue I modified Figure 6. I added...

1. ... an additional graph Figure 6 A that shows the medians (with an lower to upper quartile range ribbon) of the graves per year and region.

2. ... an indication of the mean of this medians for each 200 years time window in Figure 6 B. The square markers show, on how much data the correlation observations are based.

3. ... an indication of the p-Values in Figure 6 B to show which correlations are significant.

### 2019-06-29 article/01_Introduction.Rmd
definition of drift
proofreading

*Evolutionary archaeology*, the branch of archaeology that explicitly adopted Cultural evolution theory, established an important distinction of cultural behaviour in *function* and *style* [@dunnell1978style] which is relevant for burial traditions. Some types of human behaviour increase the chances of survival, like specific hunting techniques, superior tools or sowing more fruitful crops. Without resorting to the inappropriate simplifications of Evolutionism, their prevalence generally happens to increase over time. On the other hand, some behaviours are not related to survival and reproductive capacity, like e.g. ceramic decorations, traditional costume fashion or jewellery design. Despite or because their seemingly subordinate importance, these characteristics often pass through a breathtakingly fast development. Burial traditions can be described as a special form of these *stylistic traits*. Special because of their strong dependence on religious and spiritual beliefs and their connection with the emotionally stirring topic of death, but still subject to sometimes rapid fashion changes.

@@

@dunnell1978style postulates that the distribution and emergence of stylistic cultural traits behave randomly on a population scale level. The stochastic processes that are at work are an important object of investigation and discussed as *Neutral Transmission* or *Stylistic variability* [@bentleyRandomDriftCulture2004; @lipoNeutralityStyleDrift2001; @neimanStylisticVariationEvolutionary1995; @porcicSimulatingCulturalTransmission2014; @porcicExploringEffectsAssemblage2015; @rogersNaturalSelectionCultural2008; @shennanCeramicStyleChange2001]. @neimanStylisticVariationEvolutionary1995 takes a closer look at these processes and develops a simple framework for the simulation of non-functional idea propagation. He identifies the essential phenomena *drift* (i.e. in case of multiple, equivalent traits in a population, a single one will dominate over time due to random effects) and *flow* (i.e. intergroup transmission) in population genetics to be also relevant for cultural traits in archaeology. Furthermore, he implements a measure of cultural distance to apply his theoretical insights to real-world data. This workflow was an important methodological inspiration for the burial rites data examination in this paper.

>>>>

*Evolutionary archaeology*, the branch of archaeology that explicitly adopted Cultural evolution theory, established an important distinction of cultural behaviour in *function* and *style* [@dunnell1978style] which is relevant for burial traditions. Some types of human behaviour increase the chances of survival, like specific hunting techniques, superior tools or sowing more fruitful crops. Without resorting to the inappropriate simplifications of *Evolutionism*, their prevalence generally happens to increase over time. On the other hand, some behaviours are not related to survival and reproductive capacity, such as ceramic decorations, traditional costume fashion or jewellery design. Despite or because their seemingly subordinate importance, these characteristics often pass through a breathtakingly fast development. Burial traditions can be described as a special form of these *stylistic traits*. Special because of their strong dependence on religious and spiritual beliefs and their connection with the emotionally stirring topic of death, but still subject to sometimes rapid fashion changes.

@@

@dunnell1978style postulates that the distribution and emergence of stylistic cultural traits behave randomly on a population scale level. The stochastic processes that are at work are an important object of investigation and discussed as *Neutral Transmission* or *Stylistic variability* [@bentleyRandomDriftCulture2004; @lipoNeutralityStyleDrift2001; @neimanStylisticVariationEvolutionary1995; @porcicSimulatingCulturalTransmission2014; @porcicExploringEffectsAssemblage2015; @rogersNaturalSelectionCultural2008; @shennanCeramicStyleChange2001]. @neimanStylisticVariationEvolutionary1995 takes a closer look at these processes and develops a simple framework for the simulation of non-functional idea propagation. He identifies the essential phenomena *drift* (i.e. in case of multiple, equivalent traits in a population with no mutation or innovation, a single trait will dominate over time due to chance alone) and *flow* (i.e. intergroup transmission) in population genetics to be also relevant for cultural traits in archaeology. Furthermore, he implements a measure of cultural distance to apply his theoretical insights to real-world data. This workflow was an important methodological inspiration for the burial rites data examination in this paper.

### 2019-06-29 article/02_Materials_and_Methods.Rmd
proofreading

The code for this paper is written in the statistical programming language R [@rcoreteamLanguageEnvironmentStatistical2018] and fully reproducible [@marwickPackagingData2018]. It is available in a public Repository on GitHub (\url{https://github.com/nevrome/cultrans.bronzeageburials.article2019}) and with a permanent digital object identifier (DOI) on the Open Science Framework platform (\url{http://dx.doi.org/10.17605/OSF.IO/B6NP2}). A more in-depth description and discussion of the data preparation steps can be found in the master thesis this article is based on [@schmidComputerbasiertesCulturalEvolution2018].

@@

Regarding the variables burial type and burial construction in the grave dataset, the distribution conditions shown in Table \ref{tab:gprcrosstab} apply.

>>>>

The code for this paper is written in the statistical programming language R [@rcoreteamLanguageEnvironmentStatistical2018] and fully reproducible [@marwickPackagingData2018]. It is available in a public repository on GitHub (\url{https://github.com/nevrome/cultrans.bronzeageburials.article2019}) and with a permanent digital object identifier (DOI) on the Open Science Framework platform (\url{http://dx.doi.org/10.17605/OSF.IO/B6NP2}). A more in-depth description and discussion of the data preparation steps can be found in the master thesis this article is based on [@schmidComputerbasiertesCulturalEvolution2018].

@@

Regarding the variables *burial type* and *burial construction* in the grave dataset, the distribution conditions shown in Table \ref{tab:gprcrosstab} apply.
